Individual and Differential Harm in Redistricting

### 2023-06-02 paper/harm-short.Rmd
version without boston

---
title: Individual and Differential Harm in Redistricting
authors:
  - name: Anonymized Author(s)
    affiliation: Anonymized Institution
abstract: |
  Social scientists have developed dozens of measures for assessing partisan bias in redistricting.
  But these measures are not easily adapted to other groups, 
  including groups defined by race, class, or geography.
  Nor are they applicable to single- or no-party contexts, such as local redistricting.
  To overcome these limitations, we propose a unified framework of harm for evaluating 
  the impacts of a districting plan on individual voters and the groups to which they belong. 
  We consider a voter harmed if their chosen candidate is not elected under the current plan, 
  but would be under a different plan.
  Harm improves on existing measures by both focusing on the choices of individual voters 
  and directly incorporating counterfactual plans.
  We discuss strategies for estimating harm, and demonstrate the utility of our framework 
  through analyses of partisan gerrymandering in New Jersey, voting rights 
  litigation in Alabama, and racial dynamics of Boston City Council elections.
keywords:
  - representation
  - redistricting
  - voting rights
  - individual harm
bibliography: references.bib
biblio-style: apsr
output:
  rticles::arxiv_article:
    keep_tex: true
    includes:
      in_header: "header.tex"
    latex_engine: pdflatex
fontsize: 12pt
editor_options:
  markdown:
    wrap: sentence
---

@@

Harm and differential harm are more than a useful conceptual unification---they are practically useful in a wide variety of redistricting analyses.
A brief study of partisan gerrymandering of New Jersey congressional districts in Section \ref{sec:nj} shows how existing gerrymandering metrics can make incorrect conclusions about the magnitude and even the direction of partisan gerrymanders.
Section \ref{sec:alabama}, studies voting rights litigation for Alabama congressional districts, and uses differential harm to separate the racial and partisan effects of the enacted districting plan.
Section \ref{sec:boston} demonstrates the application of our framework to local redistricting, and quantifies the benefits of race-aware city council district boundaries to Black voters in the city of Boston.
Section \ref{sec:discussion} concludes.

@@

A second limitation is that $h_i(v,q,\tilde q)$ ignores any preferences a voter may have between the candidates that they did not vote for.
For example, a voter may consider candidates A and B nearly interchangeable, but may strongly dislike candidate C.
If the voter ultimately voted for candidate A, this harm measure makes no distinction between candidate B or candidate C winning the voter's district.
This reflects the simplistic nature of the cost function implicitly assumed in our proposed measure.
The problem could be remedied by replacing the indicator functions in the measure with other cost functions that consider ordinal preferences.
Ballot-level ranked-choice vote data, or candidate ideology scores, could be used to estimate such a cost function.
Here, for simplicity and to avoid making further assumptions, we focus on the indicator-based cost function.
Section \ref{sec:boston} further discusses the implications of this particular specification in the context of primary elections.

@@

Election results present a different set of modeling challenges, given the large number of factors that influence election outcomes and the general unpredictability of future events.
We limit our discussion here to the case of estimating harm and differential partisan harm in a two-party setting.
Estimating differential racial harm in the two-party setting can be done by first generating ecological inferences or survey estimates of party preference by race, and then applying methods analogous to those outlined here.
Multiparty or primary elections pose additional modeling challenges.
In Section \ref{sec:boston} we consider a spatial-racial model for primary elections with multiple candidates, to illustrate the application of our framework to these more complicated cases.

@@

In both these cases, the proposed harm measures quantify the extent to which a challenged districting plan differentially harms minority voters.
They account for geographic sorting, political cohesion, and the individual preferences of every voter.
Sections \ref{sec:alabama} and \ref{sec:boston} demonstrate this approach in practice.

@@

# Application to Primary Elections in Boston {#sec:boston}

The proposed harm measures are particularly easy to calculate and interpret in two-party elections, since the preferences of each kind of partisan are directly observable as partisan vote shares.
However, the same framework can be fruitfully applied to more complex settings, providing insights that go beyond those available through traditional methods.

A major question in voting rights law and research is the extent to which voting must be racially polarized before race should enter into the redistricting equation and statutory protections like the VRA should kick in [@elmendorf2016racially].
This question is particularly acute in cities, where general elections are characterized by Democratic dominance and low levels of racial polarized voting.
When the locus of political conflict shifts to primary elections, racial patterns are harder to observe and generalize.
Thus there remains uncertainty about the extent that race is a factor in primary elections, and consequently the role it should play in designing urban legislative maps.

```{r boston-map, fig.width=6.75, fig.height=3.375}
#| fig.cap: |
#|      Racial composition of the city of Boston (left)
#|      and 2010 city council districts (right).
#|      Neighborhood boundaries are overlaid in black on the
#|      left map and in white on the right map.
include_graphics("figures/boston_map.pdf")
```

The city of Boston is governed by a mayor and a nonpartisan 13-member city council.
Nine of the councilors are elected from contiguous city council districts, and the remaining four councilors are elected at-large, with all councilors serving two-year terms.
Council district boundaries are redrawn every ten years following the decennial census, and must be equipopulous, be compact, and preserve neighborhood and community boundaries.[^11]
This section will study the extent to which the current district boundaries differentially harm voters of different racial groups, in comparison to a completely race-neutral districting process.
Figure \ref{fig:boston-map} shows the racial composition of the city as well as the boundaries of the 2010 city council districts and the city's official neighborhoods.

[^11]: City of Boston charter.
    Available at <https://www.boston.gov/sites/default/files/embed/b/boston-city-charter-2007.pdf>.

## Approach

To apply our proposed framework to study the council district boundaries, we must be able to predict individual voting preferences under counterfactual districting plans.
This is more difficult than in both New Jersey and Alabama, for two reasons.
First, the city council is officially nonpartisan, and city politics are dominated by the Democratic Party, so voting choices are more complex than a simple partisan preference.
To estimate election outcomes, we must map observed support for particular candidates in one election onto counterfactual support for a hypothetical slate in another election.
Second, as in Alabama, we aim to examine the racial impact of the district system, and so must estimate these voting preferences by race.
Even if we could, for instance, divide candidates into a more "liberal" and a more "conservative" group, the vote counts for each group by precinct will not alone show how different racial groups voted.

```{r boston-factors, fig.cap="", fig.width=7, fig.height=3.5}
#| fig.cap: |
#|      Candidate factor loadings, with approximate 80\% credible intervals shown as
#|      ellipses (left), and geographic variation in factor values by race (right).
#|      Neighborhood boundaries overlaid in black.
#|      Factor values in the Black second-dimension map are truncated so that variation in the other maps is visible.
include_graphics("figures/boston_factors.pdf")
```

To solve both of these problems, we develop a *latent factor ecological inference model*, which simultaneously estimates (a) turnout by precinct and race, (b) candidate positions in a latent two-dimensional political space,[^12] and (c) voter preferences in this space by precinct and race, using precinct-level demographic data and vote totals from the March 2020 presidential primary and September 2021 mayoral preliminary elections.
The turnout component of the model is identical to that used for Alabama.
The core of the voter choice model is the following specification for $\pi_{jk}$, the preference of precinct $j$ for candidate $k$ on the logit scale:[^13] \begin{align*}
    \pi_{jk} &= \underbrace{\beta^{(s)}_k}_{\substack{\text{citywide support}\\\text{for candidate}}} +
    F_{j}^\top(\overbrace{P_j}^{\mathclap{\substack{\text{diagonal matrix of}\\\text{precinct racial proportions}}}}
    \underbrace{\turn_j}_{\mathclap{\substack{\text{turnout rate parameter}\\\text{from turnout model}}}}
    )^\top\vb L_k.
\end{align*}

[^12]: Exploratory analyses found that three latent dimensions did not provide a substantially better fit to the data.

[^13]: For simplicity of presentation, the parameterization shown here has been altered slightly from that detailed in the appendix, though the ultimate model can be made identical after an adjustment to the prior on $\vb L_k$.

Here, $F_j$ is a matrix of preferences by race and latent dimension within precinct $j$, so $F_{j}^\top(P_j\turn_j)^\top$ is a vector with elements containing the preferences of the precinct's voters in each latent dimension, after marginalizing out race.
The vector $\vb L_k$ contains the loading of candidate $k$ onto the latent space: When $\vb L_k$ and $F_{j}$ are nearby in latent space, then their product will be large, and the precinct's support for the candidate, $\pi_{jk}$, will be large.

Figure \ref{fig:boston-factors} shows the key outputs from this model, and Appendix \ref{app:lfei} explains the model in more detail.
As expected one dimension (the first) primarily aligns with a traditional liberal-conservative spectrum, with progressives like Bernie Sanders, Elizabeth Warren, and current Boston mayor Michelle Wu on the left, and centrists like Joseph R. Biden and Wu's opponent, Annissa Essaibi-George, on the right.
The second dimension appears to capture racial differences in support---as the factor value maps in Figure \ref{fig:boston-factors} show, minority-majority and White areas of the city take opposite values on the second dimension.

Like all ecological inference models, ours relies on several unverifiable assumptions.
The latent factor component imposes additional assumptions about the structure of voting in Boston elections.
As Appendix \ref{app:lfei} discusses, these assumptions are almost certainly not *strictly* met.
However, the model is a useful simplification of reality that captures the major electoral dynamics relevant to our question of interest.
For example, using the estimated latent positions and support levels for Michelle Wu and Annissa Essaibi-George, the two mayoral candidates who advanced to the general election, we predict the precinct-level general election turnout and vote shares with considerable accuracy (Spearman correlation $\approx 0.85$, see Figure \ref{fig:boston-gen-pred} in Appendix \ref{app:lfei}).
The accuracy of these validation predictions tracks with previous work using factor models to predict runoff outcomes after a first-stage election [@kamakura2006modeling].
We stress that the latent factor ecological inference model is not necessary for this kind of analysis; it is one out of several approaches which could provide estimated vote choices for individuals.
Local survey data with ideological preferences and racial information would be a sufficient alternative.

## Results

As Figure \ref{fig:boston-factors} shows, primary voting preferences in Boston are polarized by race, especially for the second latent dimension.
But there is significant variation within racial groups as well, such as the split between more conservative White voters in South Boston and more liberal White voters in Jamaica Plain (the first latent dimension appears to capture a progressive-moderate divide).
It is certainly possible, then, that the particular way districts are drawn will have an effect on the ability of different racial groups to elect their candidate of choice.

We consider two different scenarios: one in which all of the council races are contested along the first latent dimension, and one in which the races are contested along the second.
Specifically, we draw voter choices between two candidates from the posterior predictive distribution of our model, setting the candidates' loadings to --1 and 1 on the chosen dimension and 0 on the other dimension.
We draw the overall level of support from a $\Beta(15, 15)$ distribution, which means that we will study election scenarios where the overall vote share for each bloc of candidates ranges from roughly 30 to 70 percent.
The posterior predictive draws yield counts of voters for each candidate in each precinct by race, from which we can calculate winners at the district level, under any configuration of districts.

Our hypothetical elections involve only two candidates, to simplify the challenges of measuring harm with multiple candidates, over which voters may have different ranked preferences.
However, our use of latent factors provides one way to extend our harm framework to the multiple-candidate setting.
One could weight voter harm by the distance in latent space between each voter's preferred candidate and the elected candidate, thus dropping the implicit assumption in the definition of harm that all candidates the voter did not vote for are interchangeable.

We sample 2,000 districting plans across two independent runs of the SMC algorithm of @mccartan2020, setting constraints to favor compact plans that respect neighborhood boundaries.
Since we do not provide any racial information to the algorithm, this set of plans forms a good approximation of a race-neutral districting process.

For the first scenario, where council races are contested along the ideological axis, we estimate that the enacted plan is quite fair, differentially harming White voters at a rate of only 0.79% compared to the race-neutral baseline.
For the second scenario, where the contests are oriented along the racial dimension, we estimate that the enacted plan differentially harms White voters at a rate of 4.1%, compared to the race-neutral baseline.

These two harm estimates reflect the harm of the current boundaries compared to a race-neutral baseline.
The more policy-relevant direction might be in evaluating a change from the status quo.
What is the harm of switching to race-neutral districting from the current race-aware boundaries?
We can flip the above setup and compute harm by averaging over the simulated plans with the enacted plan as the counterfactual.
For this reverse comparison, along the first dimension, we estimate that a race-neutral baseline would differentially harm Black voters at a rate of 3.1% compared to the enacted plan Along the second dimension, the differential harm to Black voters would be 1.8%.

Taken together, these findings suggest that despite the dominance of the Democratic party in city politics, and the shared preferences between certain groups of Black and White voters, race-aware districting still serves an important role in empowering Black voters across the city to elect their chosen candidates.
A switch to race-neutral boundaries would have a material impact on Black voting power.
Further study could focus on the racial impact of the four at-large members in comparison to a system with 13 districts and no at-large members.

@@

Compared to existing alternatives, differential harm more accurately identifies partisan gerrymanders---both in sign and in existence.
The exact same approach applies beyond partisanship and allows for the calculation of differential harms by race and other protected groups.
By centering the framework on individual voters' choices, we sidestep the need to blindly associate voters' preferences with their group memberships.
Analysts can model vote choice separately, estimate the harm to each individual, and then aggregate by any group of interest.
As Section \ref{sec:nj}--\ref{sec:boston} showed, this opens up space for more nuanced analyses that can extend to the local level as well.

@@

# Latent Factor Ecological Inference Model for Boston Primary Elections {#app:lfei}

To estimate voter preferences by race and precinct in Boston, we use 2020 U.S.
Census demographic data and election returns from the March 2020 presidential primary and the September 2021 mayoral preliminary elections.
Our ecological inference model used to produce the estimates expands on the model in Appendix \ref{app:alei}, jointly estimating both turnout and vote choice in the two elections.
However, it estimates voter preferences in a two-dimensional latent space and also estimates candidates' positions in this space.

Continuing the notation from above, for precinct $j$ in election $v$, let $\vap_j$ denote the voting-age population, $n_{jv}$ denote the total number of votes cast for major-party candidates, $s_{jk}$ denote the vote share for candidate $k$ (who implicitly belongs to a particular election $v_k$), and $p_{jr}$ denote the fraction of the voting-age population belonging to racial group $r$, where here $r\in\{\text{White},\text{Black},\text{Hispanic},\text{Other}\}$ ("Other" consisting primarily of Asian voters in Boston).
Our main parameters of interest are $\turn_{jr}$, the baseline turnout by race and precinct, $F_{jqr}$, the preferences in latent dimension $q$ by race and precinct, and $L_{kq}$, the loading of each candidate on latent dimension $q$.
Here $q\in\{1,2\}$, since preliminary analyses found no improved fit when expanding to three dimensions.

We model the observed turnout $n_{jv}$ identically as in Alabama, above: \begin{align*}
    n_{jv} \mid \turn_{jr}, \beta^{(t)}_v
        &\sim \Binlog\qty(\vap_j, \logit(\turn_j^\top\vb p_j) + \beta^{(t)}_v) \\
    \logit(\turn_j) \mid \mu^{(t)}, \Sigma^{(t)}
        &\sim \Norm(\mu^{(t)}, \Sigma^{(t)}).
\end{align*}

Because of the latent factor model, we do not require that vote shares by race and precinct, when tallied by race, necessarily equal the observed vote share in the precinct.
As a result we choose to model the vote shares directly, rather than the vote counts, which is more computationally efficient.
Because each precinct has dozens or hundreds of voters, a Normal approximation to the Binomial likelihood is appropriate, and so we model the vote shares $s_{jk}$ as \begin{align*}
    \logit(s_{jk}) \mid \pi_{jk}, \eps &\sim \Norm(\pi_{jk}, \frac{\eps}{1+n_{jv_k}}),
\end{align*} where $\pi_{jk}$ is the preference of precinct $j$ for candidate $k$ on the logit scale, as detailed below.
The Normal approximation variance is inversely proportional to the total number of votes in the precinct (the additional 1 is to avoid divide-by-zero errors).
The multiplicative factor $\eps$ captures the additional variance that isn't accounted for by the latent factors.

The logit candidate preference itself is modeled as \begin{align*}
    \pi_{jk} &= \beta^{(s)}_k + \alpha_k
        F_{j}^\top(\turn_j\circ \vb p_j)^\top\mathrm{diag}(\vb*\rho)\vb L_k  \\
    F_{jq} \mid \Sigma^{(p)} &\sim \Norm(0, \Sigma^{(p)}).
\end{align*} Note that $F_j$ is a $q\times r$ matrix of preferences by race and dimension within precinct $j$, and $\turn_j\circ \vb p_j$ is a vector of the percentages of the total VAP in the precinct that belongs to voters of each race, so $F_{j}^\top(\turn_j\circ \vb p_j)$ is a vector with $q$ elements containing the preferences by dimension of the precinct's voters, after marginalizing out race.
The matrix $\mathrm{diag}(\vb*\rho)$ captures the relative scale or importance of each dimension to the overall outcomes.
The $\vb*\alpha$ parameters are "fudge factors" that adjust the scale of the deviations from each candidate's average support.

We give $\vb*\beta^{(t)}$ an i.i.d. $\Norm(0, 0.5^2)$ prior,and $\vb*\beta^{(s)}$ an i.i.d. $\Norm(0, 4.0^2)$ prior, and $\logit(\mu^{(t)})$ an i.i.d. $\Beta(10, 30)$ prior.
We set priors independently on the variance and correlation components of $\Sigma^{(t)}$ and $\Sigma^{(p)}$.
The former has $\Gam(1.5, 1.5/0.25)$ and $\Gam(1.5, 1.5/1.0)$ priors for turnout and support, while the latter have LKJ priors with shape parameters 4.0 and 20.0.
A larger LKJ shape parameter places more weight on correlation matrices closer to diagonal, i.e., it regularizes away from high correlations.
The scale parameters $\vb*\rho$ have i.i.d. $\Gam(1.5, 1.5/4.0)$ priors, the fudge factors $\vb*\alpha$ have i.i.d. $\Gam(5.0, 5.0)$ priors, and $\eps$ has an $\Expo(1/4.0)$ prior.
Overall, these priors weakly identify the location and scale of the parameters.
Several reparametrizations were made for computational efficiency.
The full Stan code for the model is available as part of the replication code.

Like all probabilistic factor models, ours is unidentifiable due to the rotation, location, and scale invariance of the factors in the likelihood.
To identify the location and scale, we place $\Norm(0,1)$ priors on most components of $\vb L$.
To identify the rotation, we place more informative priors on the loadings of certain candidates.
This was accomplished through somewhat of an interactive process, where a preliminary principal components decomposition of the precinct vote shares helped suggest the latent factors that might be found by the model.
We chose to align the first dimension with the Sanders-Biden axis, and chose to make the second dimension roughly orthogonal along the Janey-Buttigieg axis.
This translated to a prior of $\Norm((-1, 0), (0.1^2, 0.25^2))$ for Sanders, $\Norm((1, 0), (1.0^2, 0.5^2))$ for Biden, and $\Norm((-0.25, -1), (0.5^2, 2^2))$ for Janey.
We additionally set the prior correlation between Sanders' and Biden's loadings to be $(-0.5, 0.8)$.

The estimated loading matrix $\vb L$ is visualized on the left of Figure \ref{fig:boston-factors}, and $F$ is visualized for Black and White voters on the right.

```{r boston-gen-pred, fig.width=6.5}
#| fig.cap: |
#|      Model predictions of general election vote shares and turnout, 
#|      compared to actual general election results.
#|      Red reference line indicates perfect agreement.
include_graphics("figures/boston_gen_pred.pdf")
```

Like all ecological inference models, this one makes several assumptions, verifiable and unverifiable.
We assume no spatial correlation in residuals, when some is almost certainly present.
We place distributional assumptions on how voter preferences by races vary across precincts (importantly, that these preferences are centered around a common location).
And we of course impose a latent two-dimensional factor structure, which must always be a major simplification of reality.

To test the suitability of the model for representing the outcomes of hypothetical future elections, we use the model fitted to the primary elections to predict the outcome of the mayoral general election in November 2021.
Michelle Wu and Annissa Essaibi-George advanced to the general from the preliminary election.
We fix the model's loading matrix $\vb L$ to the posterior mean of their estimated loadings.
We draw $s_{jk}$ from a $\Beta(20\bar s, 20(1-\bar s))$ distribution, where $\bar s$ is each candidate's share of the two-candidate vote from the preliminary election (59.7% Wu -- 40.3% Essaibi George).

The model's predictions (posterior medians) for the turnout and vote share in each precinct are shown in Figure \ref{fig:boston-gen-pred}.
Considering the differences between a primary and general election, and the additional campaign time elapsed between the elections, the predictions are remarkably accurate.
The predicted turnout has a Spearman correlation of 0.84 with the actual turnout, and the predicted and actual vote shares have a correlation 0.87.
The turnout and support levels are also in relatively close agreement.

>>>>

---
title: Individual and Differential Harm in Redistricting
authors:
  - name: Cory McCartan
    thanks: |
      To whom correspondence should be addressed.
      The authors thank Steve Ansolabehere, Matt Blackwell, Kosuke Imai, Gary King, 
      Shiro Kuriwaki, Tyler Simko, Nick Stephanopolous, and 
      participants of the Fall 2021 Harvard American Politics Research Workshop,
      Fall 2022 Harvard Applied Statistics Workshop, 
      and Imai Research Group for helpful comments and discussion.
      Replication code and data are available at <https://github.com/CoryMcCartan/harm-redistricting>.
    department: Ph.D. Candidate, Department of Statistics
    affiliation: Harvard University
    location: 1 Oxford St, Cambridge, MA
    email: cmccartan@fas.harvard.edu
  - name: Christopher T. Kenny
    department: Ph.D. Candidate, Department of Government
    affiliation: Harvard University
    location: 1737 Cambridge St., Cambridge, MA
    email: christopherkenny@fas.harvard.edu
abstract: |
  Social scientists have developed dozens of measures for assessing partisan bias in redistricting.
  But these measures are not easily adapted to other groups, 
  including groups defined by race, class, or geography.
  Nor are they applicable to single- or no-party contexts, such as local redistricting.
  To overcome these limitations, we propose a unified framework of harm for evaluating 
  the impacts of a districting plan on individual voters and the groups to which they belong. 
  We consider a voter harmed if their chosen candidate is not elected under the current plan, 
  but would be under a different plan.
  Harm improves on existing measures by both focusing on the choices of individual voters 
  and directly incorporating counterfactual plans.
  We discuss strategies for estimating harm, and demonstrate the utility of our framework 
  through analyses of partisan gerrymandering in New Jersey and voting rights 
  litigation in Alabama.
keywords:
  - representation
  - redistricting
  - voting rights
  - individual harm
bibliography: references.bib
biblio-style: apsr
output:
  rticles::arxiv_article:
    keep_tex: true
    includes:
      in_header: "header.tex"
    latex_engine: pdflatex
editor_options:
  markdown:
    wrap: sentence
---

@@

Harm and differential harm are more than a useful conceptual unification---they are practically useful in a wide variety of redistricting analyses.
A brief study of partisan gerrymandering of New Jersey congressional districts in Section \ref{sec:nj} shows how existing gerrymandering metrics can make incorrect conclusions about the magnitude and even the direction of partisan gerrymanders.
Section \ref{sec:alabama}, studies voting rights litigation for Alabama congressional districts, and uses differential harm to separate the racial and partisan effects of the enacted districting plan.
Section \ref{sec:discussion} concludes.

@@

A second limitation is that $h_i(v,q,\tilde q)$ ignores any preferences a voter may have between the candidates that they did not vote for.
For example, a voter may consider candidates A and B nearly interchangeable, but may strongly dislike candidate C.
If the voter ultimately voted for candidate A, this harm measure makes no distinction between candidate B or candidate C winning the voter's district.
This reflects the simplistic nature of the cost function implicitly assumed in our proposed measure.
The problem could be remedied by replacing the indicator functions in the measure with other cost functions that consider ordinal preferences.
Ballot-level ranked-choice vote data, or candidate ideology scores, could be used to estimate such a cost function.
Here, for simplicity and to avoid making further assumptions, we focus on the indicator-based cost function.

@@

Election results present a different set of modeling challenges, given the large number of factors that influence election outcomes and the general unpredictability of future events.
We limit our discussion here to the case of estimating harm and differential partisan harm in a two-party setting.
Estimating differential racial harm in the two-party setting can be done by first generating ecological inferences or survey estimates of party preference by race, and then applying methods analogous to those outlined here.
Multiparty or primary elections pose additional modeling challenges.

@@

In both these cases, the proposed harm measures quantify the extent to which a challenged districting plan differentially harms minority voters.
They account for geographic sorting, political cohesion, and the individual preferences of every voter.
Sections \ref{sec:alabama} demonstrates this approach in practice.

@@

@@

Compared to existing alternatives, differential harm more accurately identifies partisan gerrymanders---both in sign and in existence.
The exact same approach applies beyond partisanship and allows for the calculation of differential harms by race and other protected groups.
By centering the framework on individual voters' choices, we sidestep the need to blindly associate voters' preferences with their group memberships.
Analysts can model vote choice separately, estimate the harm to each individual, and then aggregate by any group of interest.
As Section \ref{sec:nj} and \ref{sec:alabama} showed, this opens up space for more nuanced analyses that can extend to the local level as well.

@@
